Objective-First Nanophotonic Design Plan

### 2012-11-22 plan/plan.tex
Got the plan going, covered adjoint method.

\chapter{Introduction}
\section{Problem statement}
Our goal is to create a software package 
    to enable the design of nanophotonic devices.
Presently, almost all nanophotonic devices are designed
    by guessing what a good design might look like,
    based on intuition and experience,
    and then optimizing it by trial-and-error.
In contrast, we want to enable the \emph{inverse design}
    of nanophotonic devices;
    that is, to design devices by simply describing the 
    desired performance that it must achieve.

@@

\section{Capabilities of our software}\label{capabilities of our software}
We have implemented various 
    \emph{optimization paradigms} and \emph{structure parameterizations}
    which a user can arbitrarily combine in order to 
    to solve \eq{the plan (multi-mode)}.

@@

\subsection{Definition of physics residual}
The physics residual corresponds to the electromagnetic wave equation
    \E  {wave equation in E}
        {(\curl \mu^{-1} \curl - \omega^2 \epsilon) E = -i \omega J}
    which is described as $A_i(z)x_i - b_i(z)$ via
    \BI $\curl \mu^{-1} \curl - \omega^2 \epsilon \to A_i(z)$,
    \I  $\epsilon \to z$,
    \I  $E \to x_i$, and
    \I  $-i \omega J \to b_i(z)$. \EI

\subsection{Bi-affine property of the physics residual}
Critically, \eq{wave equation in E} is not only linear in $E$,
    but is also affine in $\epsilon$.
This allows us to form the extremely useful relationship,
    \E  {bi-affine property}
        {A_i(z)x_i - b_i(z) = B(x_i)z - d(x_i) = 0,}
    where $B(x_i) \in \comps^{n \times n}$ and $d(x_i) \in \comps^n$.

@@

\section{General strategy to solve the problem}

@@

    \rput[l](4.3,3){$z_\text{final}$}
    \psline{<-}(4,3)(2.4,3)
\end{pspicture}
\caption{Basic layout of our software.
        At the most basic level, the chosen optimization paradigm 
            accepts an initial structure $z_\text{guess}$
            and returns a final, optimized structure $z_\text{final}$.
        This is accomplished by repeatedly passing quadratic functions $Q(z)$ 
            to the structure parameterization,
            which returns an updated structure $z$
            which minimizes $Q(z) + g(z)$.}
\label{fig:strategy}
\end{center} \end{figure}

\subsection{Field update}
Give the \emph{specific} interfaces here. 
What exactly gets input and what does each update output?
\subsection{Structure update}

>>>>

\chapter{Introduction}
\section{Problem statement}
Our goal is to create a software package 
    to enable the design of nanophotonic devices.
Specifically, our software produces designs for
    \emph{linear} nanophotonics devices based on
    \emph{desired coupling strengths} between various optical ports.

% Presently, almost all nanophotonic devices are designed
%     by guessing what a good design might look like,
%     based on intuition and experience,
%     and then optimizing it by trial-and-error.
% In contrast, we want to enable the \emph{inverse design}
%     of nanophotonic devices;
%     that is, to design devices by simply describing the 
%     desired performance that it must achieve.

@@

\section{Structure of our software}
In reality, our software package consists of various modules
    which are designed to be maximally orthogonal.
The various modules are illustrated in \fig{structure}.

\begin{figure}[ht]\begin{center}
\begin{pspicture}(6,2.4)(-6,-2)
\psset{gridcolor=green, subgridcolor=yellow}
    \let\psgrid\relax

    \rput[t](-1.5,1){physics}
    \psframe(-0.2,0.4)(-2.8,1.4)

    \rput[t](1.5,1){simulation}
    \psframe(0.2,0.4)(2.8,1.4)

    \psline[linestyle=dashed](-3.2,0.2)(3.2,0.2)
    \rput[l](3.2,0.2){\parbox{3cm}{\center physics-to-algebra \\ translation layer}}
    \rput[t](0,-0.4){optimization}
    \psframe(-1.3,-1)(1.3,0)
\end{pspicture}
\caption{Structure of our software. 
    The physics and simulation modules provide a complete description of the problem.
    The translation module takes this description and casts it in the language of linear algebra,
        which the optimization module can then use to produce various designs.}
\label{fig:structure}
\end{center} \end{figure}

Although all modules are provided by our software, 
    the majority of the innovation lies in the optimization module.
Since the optimization module works entirely in the realm of linear algebra,
    a translation module is provided in order to cast concepts from electromagnetics into this form.
Finally, physics and simulation modules are provided 
    to allow the user to describe the design problem in physical, rather than mathematical terms.

@@

\subsection{Definition of physics residual}
The physics residual, $A_i(z)x_i - b_i(z)$, 
    corresponds to the electromagnetic wave equation
    \E  {wave equation in E}
        {(\curl \mu^{-1} \curl - \omega^2 \epsilon) E = -i \omega J}
        via
    \BI $\curl \mu^{-1} \curl - \omega^2 \epsilon \to A_i(z)$, where $\epsilon$ is a function of $z$,
    \I  $E \to x_i$, and
    \I  $-i \omega J \to b_i(z)$, typically constant with respect to $z$. \EI

\subsection{Bi-affine property of the physics residual}
Critically, \eq{wave equation in E} is not only linear in $E$,
    but is also affine in $\epsilon$.
This allows us to form the extremely useful relationship,
    \E  {bi-affine property}
        {A_i(z)x_i - b_i(z) = B_i(x_i)z - d_i(x_i) = 0,}
    where $B(x_i) \in \comps^{n \times n}$ and $d(x_i) \in \comps^n$.
   % TODO: Show what B_i and d_i are explicitly.

@@

\chapter{The optimization module}

\section{Capabilities of the module}\label{capabilities of our software}
We have implemented various 
    \emph{optimization paradigms} and \emph{structure parameterizations}
    which a user can arbitrarily combine in order to 
    to solve \eq{the plan (multi-mode)}.

\section{Structure of the module}

The module is naturally separated into two submodules, 
    the optimization paradigm (or simply, paradigm) submodule 
    and the structure parameterization (or simply, structure) submodule.
The interaction between these two submodules is illustrated in \fig{strategy}.

@@

    \rput[l](4.3,3){$z_\text{final}$}
    \psline{<-}(4,3)(2.4,3)
\end{pspicture}
\caption{Basic layout of the module.
        The optimization paradigm submodule
            accepts an initial structure $z_\text{guess}$
            and returns a final, optimized structure $z_\text{final}$.
        This is accomplished by repeatedly passing quadratic functions $Q(z)$ 
            to the structure parameterization module,
            which returns an updated structure $z$
            which minimizes $Q(z) + g(z)$.}
\label{fig:strategy}
\end{center} \end{figure}

In essense, a design is achieved via repeated calls to the structure submodule
    by the paradigm submodule,
    in which carefully chosen quadratic functions $Q(z)$ are minimized.

\section{The paradigm submodule}
\subsection{The local optimization paradigm}
The local optimization paradigm applies the adjoint method to solve \eq{rigorous problem statement}
    in a manner analogous to the steepest-descent strategy.
Specifically, we compute the derivative of the total field design objective with respect to $z$,
    \E{total derivative of field design objective}
        {\grad_z F = \sum_i^N \grad_z f_i(x_i),}
    and then pass 
    \E{}{Q(z) = \frac{1}{2}\norm{z-z_0}^2 + \kappa \grad_z F\cT (z-z_0)}
    to the structure submodule, for some $\kappa$.

The analogy with a steepest-descent strategy is apparent since 
    the global minimum of $Q(z)$ can be found via
    \E{}{\grad_z Q(z) = (z - z_0) + \kappa \grad F = 0,}
    resulting in 
    \E{}{z = z_0 - \kappa \grad_z F,}
    from which we see that the role of $\kappa$ is to determine the step-size 
    in the direction of steepest-descent.

\subsubsection{Computation of $\grad_z F$}
The derivatives of the individual field design objectives are found via
    \E{}{\frac{d}{dz} f_i(x_i) = \frac{\pd f_i}{\pd x_i} \frac{dx_i}{dz}.}
To obtain ${dx_i}/{dz}$ we first differentiate the corresponding physics residual, 
    $A_i(z) x_i - b_i(z)$,
    \E{}{A_i(z) dx_i + dA_i(z) - db_i(z) = A_i(z)dx_i + B_i(x_i)dz,}
    where $dA_i(z) - db_i(z) = B_i(x_i)dz$ 
    as a result of the bi-affine property of the physics residual \eq{bi-affine property}.

Assuming that physics is already satisfied, $A_i(z) x_i - b_i(z) = 0$,
    and that we want to keep the physics residual at 0,
    the following condition must then be satisfied,
    \E{}{A_i(z) dx_i = -B_i(x_i) dz}
    from which we obtain
    \E{}{\frac{dx}{dz} = -A_i(z)^{-1} B_i(x_i).}

Efficient calculation of $dx/dz$ is almost always impossible since $B_i(x_i) \in \comps^{n \times n}$.
At the same time, since ${\pd f_i}/{\pd x_i} \in \comps^{1 \times m}$, we instead compute $df_i(x_i)/dz$ via
    \E{}{\frac{d}{dz} f_i(x_i) = -\frac{\pd f_i}{\pd x_i} A_i(z)^{-1} B_i(x_i) = 
        -\left(A_i(z)^\cT \frac{\pd f_i}{\pd x_i}^\cT\right)^\cT B_i(x_i)}
    which requires only one solve of $A_i^\cT$ as opposed to $n$ solves of $A_i$.

\subsubsection{Computation of $\pd f_i(x_i) / \pd x_i$}
The definition of $f_i(x_i)$ as given in \eq{field design objective definition} is not differentiable 
    since any deviation away from the power constraints results in $f_i = \infty$.
In order to make the constraints differentiable, then, 
    we use the following relaxed indicator function $I^\text{rel}_+$ in place of $I_+$,
    \E{}{I^\text{rel}_+(u) = \begin{cases} 0 & u \ge 0, \\ 
                                \frac{1}{a}|u|^p & u < 0, \end{cases}}
    where $a$ is a normalization factor and $p \in (0, \infty]$. 
In order to guarantee a well-defined value of $\pd f_i(x_i) / \pd x_i$ even for $p = \infty$,
    we can let $a = \max_i f_i(x_i)$.

\subsection{The global optimization paradigm}

\section{The structure submodule}
